# Balance of Payments Table 12

using WEO data from IMF

Check with print version WESP 2025:
- slight dviation for a Western Asiam Fuel Importing Country (probably Yemen)
- feeds trough to other aggregates


**UPDATE HERE**: START AND END YEAR FOR THE TABLE (end year can be in the future, WEO has projections)
```{r}
start_year <- 2016
end_year <- 2026
```

## Imports

the usual libraries...
```{r}
library(tidyverse)
library(rsdmx) # for the API access
```

## Load the WEO dataset

BCA = indicator for Current account balance in USD
```{r}
flowref <- 'WEO'
key <- "*.BCA.A"
```

```{r}
dataset_IMF <- as.data.frame(readSDMX(providerId = 'IMF_DATA',
                                  resource = 'data',
                                  flowRef = flowref,
                                  key = key
                                  ))
```

## Cleaning

fix the palestine code in the IMF (from WBG for West Bank and Gaza to PSE)
```{r}
dataset_IMF <- dataset_IMF %>%
  mutate(
    COUNTRY = ifelse(COUNTRY == "WBG", "PSE", COUNTRY)
  )
```


WEFM classications of countries (created in file `00_WESP_classification.qmd`)
```{r}
wesp <- haven::read_dta("./data/WESP_classification.dta") 
```

```{r}
df <- dataset_IMF %>%
  select(TIME_PERIOD, COUNTRY, OBS_VALUE) %>%
  filter(TIME_PERIOD >= start_year & TIME_PERIOD <= end_year) %>%
  right_join(wesp, by = c("COUNTRY" = "ISO3")) %>%
  mutate(
    OBS_VALUE = as.numeric(OBS_VALUE)
  )
```

## Wrangling

### Developed economies

```{r}
row_developed <- df %>%
  filter(status == "DVD") %>%
  mutate(group = "**Developed Economies**") %>%
  group_by(group, TIME_PERIOD) %>%
  reframe(
    value = sum(OBS_VALUE, na.rm = TRUE), 
    value = value / 1e9 # in billions
  ) %>%
  pivot_wider(names_from = TIME_PERIOD, values_from = value) 

row_developed
```
now individual DVD countries

```{r}
row_developed_countries <- df %>%
  filter(
    status == "DVD",
    COUNTRY %in% c("USA", "JPN", "KOR") #"GBR"
    ) %>%
  mutate(group = country) %>%
  select(group, TIME_PERIOD, OBS_VALUE) %>%
  mutate(
    OBS_VALUE = as.numeric(OBS_VALUE) / 1e9 # in billions
  ) %>%
  pivot_wider(names_from = TIME_PERIOD, values_from = OBS_VALUE)
```

some groups of DVD countries

```{r}
row_developed_groups <- df %>%
  filter(
    status == "DVD",
  ) %>%
  mutate(
    group = case_when(
      dummy_EU27 == 1 ~ "EU27",
      dummy_otherEUR == 1 ~ "Other Europe",
      COUNTRY %in% c("AUS", "CAN", "NZL") ~ "Other Developed Economies",
    )
  ) %>%
  group_by(group, TIME_PERIOD) %>%
  reframe(
    value = sum(OBS_VALUE, na.rm = TRUE), 
    value = value / 1e9 # in billions
  ) %>%
  pivot_wider(names_from = TIME_PERIOD, values_from = value) %>%
  filter(!is.na(group))
```

### Economies in Transtion

```{r}
row_eit <- df %>%
  filter(status == "EIT") %>%
  mutate(group = "**Economies in Transition**") %>%
  group_by(group, TIME_PERIOD) %>%
  reframe(
    value = sum(OBS_VALUE, na.rm = TRUE), 
    value = value / 1e9 # in billions
  ) %>%
  pivot_wider(names_from = TIME_PERIOD, values_from = value)

row_eit
```


```{r}
row_eit_groups <- df %>%
  filter(status == "EIT") %>%
  mutate(
    group = case_when(
      dummy_SEE == 1 ~ "South-Eastern Europe",
      dummy_CIS == 1 ~ "Commonwealth of Independent States",
    )
  ) %>%
  group_by(group, TIME_PERIOD) %>%
  reframe(
    value = sum(OBS_VALUE, na.rm = TRUE), 
    value = value / 1e9 # in billions
  ) %>%
  pivot_wider(names_from = TIME_PERIOD, values_from = value) %>%
  filter(!is.na(group)) %>%
  # switch rows order
  arrange(desc(group))

row_eit_groups
```

## Developing economies

### All DVG
```{r}
row_developing <- df %>%
  filter(status == "DVG") %>%
  mutate(group = "**Developing Economies**") %>%
  group_by(group, TIME_PERIOD) %>%
  reframe(
    value = sum(OBS_VALUE, na.rm = TRUE), 
    value = value / 1e9 # in billions
  ) %>%
  pivot_wider(names_from = TIME_PERIOD, values_from = value)

row_developing
```


```{r}
# now the exporter / importer groups
row_developing_group <- df %>%
  filter(status == "DVG") %>%
  mutate(group = case_when(
    dummy_exporter == 1 ~ "Net Fuel Exporters",
    dummy_exporter == 0 ~ "Net Fuel Importers",
  )) %>%
  group_by(group, TIME_PERIOD) %>%
  reframe(
    value = sum(OBS_VALUE, na.rm = TRUE), 
    value = value / 1e9 # in billions
  ) %>%
  pivot_wider(names_from = TIME_PERIOD, values_from = value)

row_developing_group
```


### Subregions

#### LAC

```{r}
df_raw <- df %>%
  filter(
    status == "DVG",
    wesp_region == "Latin America and the Caribbean"
  ) 

df_raw_all <- df_raw %>% mutate(group = "*Latin America and the Caribbean*") 

df_raw_group <- df_raw %>%
  mutate(
    group = case_when(
      dummy_exporter == 1 ~ "Net Fuel Exporters",
      dummy_exporter == 0 ~ "Net Fuel Importers",
    )
  )

row_LAC <- bind_rows(df_raw_all, df_raw_group) %>%
  group_by(group, TIME_PERIOD) %>%
  reframe(
    value = sum(OBS_VALUE, na.rm = TRUE), 
    value = value / 1e9 # in billions
  ) %>%
  pivot_wider(names_from = TIME_PERIOD, values_from = value)
```

#### Africa

```{r}
df_raw <- df %>%
  filter(
    status == "DVG",
    wesp_region == "Africa"
  ) 

df_raw_all <- df_raw %>% mutate(group = "*Africa*") 

df_raw_group <- df_raw %>%
  mutate(
    group = case_when(
      dummy_exporter == 1 ~ "Net Fuel Exporters",
      dummy_exporter == 0 ~ "Net Fuel Importers",
    )
  )

row_AFR <- bind_rows(df_raw_all, df_raw_group) %>%
  group_by(group, TIME_PERIOD) %>%
  reframe(
    value = sum(OBS_VALUE, na.rm = TRUE), 
    value = value / 1e9 # in billions
  ) %>%
  pivot_wider(names_from = TIME_PERIOD, values_from = value)
```

#### Western Asia

```{r}
df_raw <- df %>%
  filter(
    status == "DVG",
    wesp_subregion == "Western Asia"
  ) 

df_raw_all <- df_raw %>% mutate(group = "*Western Asia*") 

df_raw_group <- df_raw %>%
  mutate(
    group = case_when(
      dummy_exporter == 1 ~ "Net Fuel Exporters",
      dummy_exporter == 0 ~ "Net Fuel Importers",
    )
  )

row_WASIA <- bind_rows(df_raw_all, df_raw_group) %>%
  group_by(group, TIME_PERIOD) %>%
  reframe(
    value = sum(OBS_VALUE, na.rm = TRUE), 
    value = value / 1e9 # in billions
  ) %>%
  pivot_wider(names_from = TIME_PERIOD, values_from = value)
```


#### East and Southeast Asia

```{r}
df_raw <- df %>%
  filter(
    status == "DVG",
    wesp_region == "Asia",
    wesp_subregion != "Western Asia" # the leftoer regions are the other 2
  ) 

df_raw_all <- df_raw %>% mutate(group = "*East and South Asia*") 

df_raw_group <- df_raw %>%
  mutate(
    group = case_when(
      dummy_exporter == 1 ~ "Net Fuel Exporters",
      dummy_exporter == 0 ~ "Net Fuel Importers",
    )
  )

row_EASIA <- bind_rows(df_raw_all, df_raw_group) %>%
  group_by(group, TIME_PERIOD) %>%
  reframe(
    value = sum(OBS_VALUE, na.rm = TRUE), 
    value = value / 1e9 # in billions
  ) %>%
  pivot_wider(names_from = TIME_PERIOD, values_from = value)
```

## World

```{r}
row_world <- df %>%
  mutate(group = "**World**") %>%
  group_by(group, TIME_PERIOD) %>%
  reframe(
    value = sum(OBS_VALUE, na.rm = TRUE), 
    value = value / 1e9 # in billions
  ) %>%
  pivot_wider(names_from = TIME_PERIOD, values_from = value)

row_world
```


## Combination

```{r}
df_final<-  bind_rows(
    row_developed,
    row_developed_countries,
    row_developed_groups,
    row_eit,
    row_eit_groups,
    row_developing,
    row_developing_group,
    row_AFR,
    row_EASIA,
    row_WASIA,
    row_LAC,
    row_world
  ) %>%
  select(-`NA`) %>% # remove NA artifact column
  mutate(across(-group, ~round(.x, 1))) # round to 1 decimal place
 
df_final
```

```{r}
writexl::write_xlsx(df_final, "./output/table_12_bop.xlsx")
```

