# Table 9

Table 9 releis on UNCTAD data for:
- commodity prices for selected commodities
- from https://unctadstat.unctad.org/datacentre/dataviewer/US.UCPI_A

Check with print version WESP 2025: exactly the same


**UPDATE HERE**: BASE YEAR = CURRENT YEAR
```{r}
base_year <- 2025
```


## Getting the data

Method
- uses the bulk download permalink to get all years of data
- extracts the donwload into a temporary directory and unzips
- then uses the CSV

Reason: API needs key, which is more difficult than just this

```{r}
library(tidyverse)
```

Annual Data
```{r}
url_A <- "https://unctadstat-api.unctad.org/bulkdownload/US.UCPI_A/US_UCPI_A"

file_7z_A <- tempfile(fileext = ".7z")
dir_A     <- tempfile("unctad_annual_")

dir.create(dir_A)

download.file(url_A, file_7z_A, mode = "wb")
archive::archive_extract(file_7z_A, dir = dir_A)

csv_A <- list.files(dir_A, pattern = "\\.csv$", full.names = TRUE)
df_A  <- read_csv(csv_A)

on.exit(unlink(dir_A, recursive = TRUE), add = TRUE)
```

Monthly Data (for most recent year calcs)
```{r}
url_M <- "https://unctadstat-api.unctad.org/bulkdownload/US.UCPI_M/US_UCPI_M"

file_7z_M <- tempfile(fileext = ".7z")
dir_M     <- tempfile("unctad_monthly_")

dir.create(dir_M)

download.file(url_M, file_7z_M, mode = "wb")
archive::archive_extract(file_7z_M, dir = dir_M)

csv_M <- list.files(dir_M, pattern = "\\.csv$", full.names = TRUE)
df_M  <- read_csv(csv_M)

on.exit(unlink(dir_M, recursive = TRUE), add = TRUE)
```

## Processing

```{r}
df_A_clean <- df_A %>%
  reframe(
    Year        = as.integer(Year),
    Commodity   = `Commodity Label`,
    Price_Index = as.double(`Index Base 2015`),
  ) %>%
  pivot_wider(
    names_from  = Commodity,
    values_from = Price_Index
  ) %>%
  select(
    Year, Food, `Tropical beverages`, `Vegetable oilseeds and oils`, `Agricultural raw materials`,
    `Minerals, ores and metals`, `All groups`, `All groups excl. fuels`, `Fuels`
  ) %>%
  filter(Year >= base_year-9) # last 10 years
  # filter(Year >= 2015) # from 2015 onwards

df_A_clean
```

now the monthly calculations for the most recent year, sorted into Quarters

```{r}
df_M_clean <- df_M %>%
  reframe(
    Year  = as.integer(substr(Period, 1, 4)),
    month = as.integer(substr(Period, 6, 7)),
    Commodity   = `Commodity Label`,
    Price_Index = as.double(`Index Base 2015`),
  ) %>%
  pivot_wider(
    names_from  = Commodity,
    values_from = Price_Index
  ) %>%
  filter(Year >= base_year-2) %>% # last 3 years should have quarter-wise analysis
  group_by(Year) %>%
  mutate(quarter = case_when(
    month %in% 1:3   ~ 1,
    month %in% 4:6   ~ 2,
    month %in% 7:9   ~ 3,
    month %in% 10:12 ~ 4,
    TRUE             ~ NA_real_
  )) %>%
  group_by(Year, quarter) %>%
  filter(n() == 3) %>% # filter out if quarter has less than 3 months of data
  summarise_all(mean) %>%
  reframe(
    Year = paste0(Year, " Q", quarter),
    Food, `Tropical beverages`, `Vegetable oilseeds and oils`, `Agricultural raw materials`,
    `Minerals, ores and metals`, `All groups`, `All groups excl. fuels`, `Fuels`
  ) 

df_M_clean
```

## Combining

```{r}
df_combined <- df_A_clean %>%
  mutate(Year = as.character(Year)) %>%
  bind_rows(df_M_clean) %>%
  mutate(across(-Year, ~ round(.x, 0))) 

df_combined
```

Saving
```{r}
writexl::write_xlsx(df_combined, "output/table_9_commodity_price_indices.xlsx")
```

