# Table 13

NOTE:
- updated groupings for the first 2 columns to reflect 10 year intervals
- selection of EU countries adjustable in last code block (my selection = 4 largest)
- most recent year may differ each time you compile, because of data updates

CHECK with print version
- 2024 data not comparable bcause of data updates (but very close)
- growth data for 2022, 2023 = exactly the same (aka the calcualtion should be correct)
- because of time frame updates: 10 year windwows not comparable

**UPDATE HERE**: BASE YEAR = CURRENT YEAR - 1 (most of the time, depends on data availability)
```{r}
base_year <- 2024
```

How it works (broadly):
- go to the online Table viewer at OECD stats site (http://stats.oecd.org/index.aspx?datasetcode=table1)
- Select some groupings / measurements on the left side (described at respective parts of the code)
- go to "Developer API" tab and copy the Data Query for SDMX
=> these are the urls in the code below

--- 

```{r}
library(tidyverse)
library(rsdmx)
```

## Columns: growth rates

Online Table viewer: http://stats.oecd.org/index.aspx?datasetcode=table1

Selection:
- Donor: DAC countries (dont forget the DAC EU countries box at the bottom)
- Measure: Official Development Assistance (ODA), Disbursements net
- Flow type: Disbursements, net
- Price base: Constant
- Combined unit of measure: US dollar, Millions, 2023

```{r}
url <- "https://sdmx.oecd.org/public/rest/data/OECD.DCD.FSD,DSD_DAC1@DF_DAC1,/AUS+AUT+BEL+CAN+CZE+DNK+EST+FIN+FRA+DEU+GRC+HUN+ISL+IRL+ITA+JPN+KOR+LVA+LTU+LUX+NLD+NZL+NOR+POL+PRT+SVK+SVN+ESP+SWE+CHE+GBR+USA+DAC+DACEU._Z.1010..1140.USD.Q?startPeriod=2000&dimensionAtObservation=AllDimensions"
```

```{r}
dataset_1 <- readSDMX(url)
df_1 <- as.data.frame(dataset_1)
```

Method: calculate year-on-year growth rates, then average them over the respective periods
Note: if you want to adjust the periods, do it in the `case_when` part below

```{r}
df_growth <- df_1 %>%
  select(DONOR, TIME_PERIOD, obsValue) %>%
  # calc growth rate
  arrange(TIME_PERIOD) %>%
  group_by(DONOR) %>%
  mutate(growth_rate = (obsValue / lag(obsValue) - 1) * 100) %>%
  # create groups for summing up
  mutate(group = case_when(
    TIME_PERIOD == base_year ~ as.character(base_year),
    TIME_PERIOD == (base_year - 1) ~ as.character(base_year - 1),
    TIME_PERIOD == (base_year - 2) ~ as.character(base_year - 2),
    TIME_PERIOD <= (base_year - 3) & TIME_PERIOD >= (base_year - 12) ~ paste0(base_year - 12,"-", base_year - 3),
    TIME_PERIOD <= (base_year - 13) & TIME_PERIOD >= (base_year - 22) ~ paste0(base_year - 22,"-", base_year - 13)
    )
  ) %>%
  filter(!is.na(group)) %>%
  group_by(DONOR, group) %>%
  reframe(
    growth_rate_avg = mean(growth_rate, na.rm = T),
    growth_rate_avg = round(growth_rate_avg, 1)
    ) %>%
  # pivot to make original table
  pivot_wider(names_from = group, values_from = growth_rate_avg) 

#df_growth
```


## Column: total ODA

same Table: http://stats.oecd.org/index.aspx?datasetcode=table1

Selection:
- Donor: DAC countries (dont forget the DAC EU countries box at the bottom)
- Measure: Official Development Assistance (ODA), Disbursements net
- Flow type: Disbursements, net
- Price base: Current
- Combined unit of measure: US dollar, Millions


```{r}
url <- "https://sdmx.oecd.org/public/rest/data/OECD.DCD.FSD,DSD_DAC1@DF_DAC1,/AUS+AUT+BEL+CAN+CZE+DNK+EST+FIN+FRA+DEU+GRC+HUN+ISL+IRL+ITA+JPN+KOR+LVA+LTU+LUX+NLD+NZL+NOR+POL+PRT+SVK+SVN+ESP+SWE+CHE+GBR+USA+DAC+DACEU._Z.1010..1140.USD.V?startPeriod=2020&dimensionAtObservation=AllDimensions"
```

```{r}
dataset_current <- readSDMX(url)
df_current <- as.data.frame(dataset_current)
```


```{r}
df_total <- df_current %>%
  select(DONOR, TIME_PERIOD, obsValue) %>%
  filter(TIME_PERIOD == base_year) %>%
  reframe(
    DONOR,
    TOTAL_ODA = round(obsValue / 1e3, 3)
  ) 

#df_total
```

## Column: Percentage of GNI

same table: http://stats.oecd.org/index.aspx?datasetcode=table1

Selection: 
- Donor: DAC countries (dont forget the DAC EU countries box at the bottom)
- Measure: GNI at current prices
- Flow type: Disbursements, net
- Price base: Current
- Combined unit of measure: US dollar, Millions

```{r}
url <- "https://sdmx.oecd.org/public/rest/data/OECD.DCD.FSD,DSD_DAC1@DF_DAC1,/AUS+AUT+BEL+CAN+CZE+DNK+EST+FIN+FRA+DEU+GRC+HUN+ISL+IRL+ITA+JPN+KOR+LVA+LTU+LUX+NLD+NZL+NOR+POL+PRT+SVK+SVN+ESP+SWE+CHE+GBR+USA+DAC+DACEU._Z.1..1140.USD.V?startPeriod=2015&dimensionAtObservation=AllDimensions"
```

```{r}
dataset_2 <- readSDMX(url)
df_2 <- as.data.frame(dataset_2)
```

```{r}
df_gni <- df_2 %>%
  select(DONOR, TIME_PERIOD, obsValue) %>%
  filter(TIME_PERIOD == base_year) %>%
  reframe(DONOR, GNI = obsValue) %>%
  left_join(df_total, by = "DONOR") %>%
  mutate(
    GNI = GNI/1e3,
    PCT_GNI = round( (TOTAL_ODA / GNI) * 100 , 2)
  ) %>%
  select(DONOR, PCT_GNI)

#df_gni
```

## Columns: distribution



same table, different Selection: 
- Donor: DAC countries (dont forget the DAC EU countries box at the bottom)
- Measure: 
  - Bilateral ODA
  - Multilateral ODA
    - UN Agencies
- Flow type: Disbursements, net
- Price base: Current
- Combined unit of measure: US dollar, Millions

```{r}
url <- "https://sdmx.oecd.org/public/rest/data/OECD.DCD.FSD,DSD_DAC1@DF_DAC1,/AUS+AUT+BEL+CAN+CZE+DNK+EST+FIN+FRA+DEU+GRC+HUN+ISL+IRL+ITA+JPN+KOR+LVA+LTU+LUX+NLD+NZL+NOR+POL+PRT+SVK+SVN+ESP+SWE+CHE+GBR+USA+DAC+DACEU._Z.1015+2000+2101..1140.USD.V?startPeriod=2015&dimensionAtObservation=AllDimensions"
```

```{r}
dataset_3 <- readSDMX(url)
df_3 <- as.data.frame(dataset_3)
```

```{r}
df_distribution <- df_3 %>%
  filter(TIME_PERIOD == base_year) %>%
  mutate(
    target = case_when(
      MEASURE == "2101" ~ "UN",
      MEASURE == "2000" ~ "Multilateral",
      MEASURE == "1015" ~ "Bilateral"
    )
  ) %>%
  select(DONOR, target, obsValue) %>%
  pivot_wider(names_from = target, values_from = obsValue) %>%
  # distributions
  mutate(
    total = Bilateral + Multilateral,
    multi_other = Multilateral - UN,
    pct_bilateral = round((Bilateral / total) * 100, 1),
    pct_multilateral = round((Multilateral / total) * 100, 1),
    pct_un = round((UN / total) * 100, 1),
    pct_multi_other = round((multi_other / total) * 100, 1)
  ) %>%
  select(DONOR, pct_bilateral, pct_multilateral, pct_un, pct_multi_other)

#df_distribution
```


## Final Table

Assemble all together, adjust the filter to only have certain countries in the final table

```{r}
df_table <- df_growth  %>%
  left_join(df_gni, by = "DONOR") %>% # combine with pct of GNI
  left_join(df_total, by = "DONOR") %>% # combine with total ODA for base year
  left_join(df_distribution, by = "DONOR") %>% # combine with distribution
  filter(
    DONOR %in% c(
      "DAC", "DACEU", # aggregates
      "FRA", "DEU", "ITA", "ESP", # selected EU countries
      # all other EU countries COMMENT IN IF NEEDED
      # "AUT","BEL", "CZE","DNK","EST","FIN","GRC","HUN","IRL","ISL","LTU","LUX","LVA","NLD","POL","PRT","SVK","SVN","SWE",
      # other DAC countries
      "AUS", "CAN", "JPN", "KOR", "NZL", "NOR", "CHE", "GBR", "USA"
      )
  ) %>%
  arrange(factor(DONOR, levels = c(
    "DAC", "DACEU", 
    "FRA", "DEU", "ITA", "ESP", # selected EU countries
    # all other EU countries COMMENT IN IF NEEDED
    # "AUT","BEL", "CZE","DNK","EST","FIN","GRC","HUN","IRL","ISL","LTU","LUX","LVA","NLD","POL","PRT","SVK","SVN","SWE",
    # other DAC countries
    "AUS", "CAN", "JPN", "KOR", "NZL", "NOR", "CHE", "GBR", "USA" 
    )))

df_table
```

Save it to an excel file

```{r}
writexl::write_xlsx(df_table, "output/table_13_oecd_oda.xlsx")
```


