# Table 14


NOTE: 
- most recent year may differ each time you compile, because of data updates
- added the option of In-Donor Costs (potentially relevant because of Refugees)
  - comment it out in last code block to remove it from the table
  
CHECK with 2025 WESP:
- Debt relief: slight deviation due to using the Bilateral Data and not the Total ODA data
- Other Multilaterals: calculated as differential, therefore new numbers
- Loans: used the Loans Data available in Table 2a, I was not able to reproduce the ones in the WESP table

Also: 
- Humanitarian Aid + Loans are NOT exclusively Bilateral! 
- Table2a from OECD combines disbursements from countries as well as Multilat. Org. 

**UPDATE HERE:** BASE YEAR = current year - 1 (most of the time, depends on data availability)

```{r}
base_year <- 2024
```

How it works (broadly):
- go to the online Table viewer at OECD stats site (e.g http://stats.oecd.org/index.aspx?datasetcode=table1)
- Select some groupings / measurements on the left side (described at respective parts of the code)
- go to "Developer API" tab and copy the Data Query for SDMX
=> these are the urls in the code below


---

```{r}
library(tidyverse)
library(rsdmx)
```

## Load the Data

Online table viewer: http://stats.oecd.org/index.aspx?datasetcode=table1

Selection:
- Donor: DAC countries 
- Measure: Official Development Assistance (ODA), and subcategories
 - Bilateral ODA
   - Debt relief
   - Other In-donor expenditures
   - Memo: Free Standing Technical Cooperation
  - Multilateral ODA
    - UN Agencies
    - EU Institutions
    - IDA
    - Other World Bank
    - Regional Development Banks
- Flow type: Disbursements, net
- Price base: Current prices
- Combined unit of measure: US dollar, Millions

API access via SDMX
```{r}
url <- "https://sdmx.oecd.org/public/rest/data/OECD.DCD.FSD,DSD_DAC1@DF_DAC1,/DAC._Z.1902+1800+1010+1600+2102+2103+2104+2105+1015+2000+2101..1140.USD.V?startPeriod=2015&dimensionAtObservation=AllDimensions"
```

```{r}
dataset_4 <- readSDMX(url)
df_4 <- as.data.frame(dataset_4)
```

## Cleaning 

```{r}
table_final <- df_4 %>%
  filter(TIME_PERIOD > base_year-10) %>%
  mutate(
    target = case_when(
      MEASURE == "1010" ~ "**Total ODA**",
      # Bilat
      MEASURE == "1015" ~ "*Bilateral*",
      MEASURE == "1902" ~ "Technical Cooperation",
      # MEASURE == "1310" ~ "Investment Projects",
      MEASURE == "1600" ~ "Debt relief",
      MEASURE == "1800" ~ "In Donor Costs",
      
      # multilateral
      MEASURE == "2000" ~ "*Multilateral*",
      MEASURE == "2101" ~ "UN agencies",
      MEASURE == "2102" ~ "EU Institutions",
      MEASURE == "2103" ~ "World Bank", # IDA
      MEASURE == "2104" ~ "World Bank", # IBRD, IFC
      MEASURE == "2105" ~ "Regional Development Banks"
    )
  ) %>%
  # combine WB rows
  group_by(TIME_PERIOD, target) %>%
  reframe(obsValue = sum(obsValue) / 1e3) %>%
  # pivot
  select(target, TIME_PERIOD, obsValue) %>%
  pivot_wider(names_from = TIME_PERIOD, values_from = obsValue) 

table_final 
```

### Other Mulitlaterals

```{r}
row_other_mult <- table_final %>%
  pivot_longer(-target, names_to = "TIME_PERIOD", values_to = "obsValue") %>%
  filter(target %in% c("*Multilateral*", "UN agencies", "EU Institutions", "World Bank", "Regional Development Banks")) %>%
  group_by(TIME_PERIOD) %>%
  # row wise: *Multilateral* - the rest
  reframe(obsValue = obsValue[target == "*Multilateral*"] - sum(obsValue[target != "*Multilateral*"])) %>%
  pivot_wider(names_from = TIME_PERIOD, values_from = obsValue) %>%
  mutate(target = "Other Multilaterals")
  
row_other_mult
```


### Humanitarian Aid data

NOTE: Drawn from a different Table!

Online table viewer: http://stats.oecd.org/index.aspx?datasetcode=table2a

Selection
- Donor:  DAC countries
- Recipient: All recipients
- Measure: 
  - Humanitarian aid
  - ODA Loans, net disbursements
  - Debt forgiveness
- Price base: Constant prices
- Combined unit of measure: US dollar, Millions, 2023

```{r}
url <- "https://sdmx.oecd.org/public/rest/data/OECD.DCD.FSD,DSD_DAC2@DF_DAC2A,/DAC.ALLR.216+218.USD.V?startPeriod=2015&dimensionAtObservation=AllDimensions"
```

```{r}
dataset_ha <- readSDMX(url)
df_ha <- as.data.frame(dataset_ha)
```

```{r}
rows_ha <- df_ha %>%
  mutate(target = case_when(
    # MEASURE == 212 ~ "Debt forgiveness",
    MEASURE == 216 ~ "Humanitarian Aid",
    MEASURE == 218 ~ "Loans"
  )) %>%
  group_by(TIME_PERIOD, target) %>%
  reframe(obsValue = sum(obsValue) / 1e3) %>%
  pivot_wider(names_from = TIME_PERIOD, values_from = obsValue)

rows_ha
```





### LDC data 

Online table viewer (same as for Humanitarian Aid): http://stats.oecd.org/index.aspx?datasetcode=table2a

Selection
- Donor: DAC countries
- Recipient: Least developed countries
- Measure: Official development assistance (ODA), disbursements
- Price base: Constant prices
- Combined unit of measure: US dollar, Millions, 2023

```{r}
url <- "https://sdmx.oecd.org/public/rest/data/OECD.DCD.FSD,DSD_DAC2@DF_DAC2A,/DAC.LDC.206.USD.V?startPeriod=2010&dimensionAtObservation=AllDimensions"
```

```{r}
dataset_ldc <- readSDMX(url)
df_ldc <- as.data.frame(dataset_ldc)
```

```{r}
row_ldc <- df_ldc %>%
  select(TIME_PERIOD, obsValue) %>%
  filter(TIME_PERIOD > base_year-10) %>%
  mutate(obsValue = obsValue / 1e3) %>% # in billions
  arrange(TIME_PERIOD) %>%
  pivot_wider(names_from = TIME_PERIOD, values_from = obsValue) %>%
  mutate(target = "*Memo Item*: ODA to LDC") 

row_ldc
```


## Final table

Assembling all parts of the final table, and sort according to the original order

```{r}
table_comb <- table_final %>%
  bind_rows(rows_ha) %>%
  bind_rows(row_ldc) %>%
  bind_rows(row_other_mult) %>%
  arrange(factor(target, levels = c(
    "**Total ODA**",
    "*Bilateral*",
    "Technical Cooperation",
    "Debt relief",
    "In Donor Costs",
    "Humanitarian Aid",
    "Loans",
    "*Multilateral*",
    "UN agencies",
    "EU Institutions",
    "World Bank",
    "Regional Development Banks",
    "Other Multilaterals",
    "*Memo Item*: ODA to LDC"
  ))) %>%
  mutate(across(-target, ~ round(.x, 1))) #%>%
  # FILTER FOR IN DONOR COSTS IF NEEDED
  # filter(target != "In Donor Costs")
  

table_comb
```

Save the table as excel
```{r}
writexl::write_xlsx(table_comb, "output/table_14_oecd_oda_breakdown.xlsx")
```

